Do Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates Affect Rice Sheath Blight?

### 2018-05-02 analysis/paper/paper.Rmd
Replace "nitrogen" with "N"

Sheath blight (ShB) disease (*Rhizoctonia solani* Kühn), anastamosis group 1 [(teleomorph: *Thanatephorus cucumeris* (A.B. Frank) Donk.] of rice is an economically important disease worldwide throughout tropical and temperate production areas. Under conducive conditions he disease can cause up to 50\% yield losses [@Marchetti1991]. Studies by Savary et al. [-@Savary2000] estimated normal yield losses in tropical irrigated lowland rice in Asia at 5-10\%. The disease is best managed through an integrated disease management approach which includes resistance [@Kumar2009, @Srinivasachary2011] fungicide applications where affordable [@Rush1983; @Groth2008], nitrogen management [@Slaton2003; @Castilla1996; @Tang2007] and water management [@Castilla1996].

@@

The plot design was a split-plot design with four replicates where irrigation was the main plot and nitrogen (N) rate was the split-plot.

@@

The split-plot treatment rates of nitrogen were N0 (no nitrogen supply), N100 (100 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation) and N120 (120 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation) (Table \@ref(tab:N-rates)).

@@

Split-plot N rates differed from the 2015 rates with only two nitrogen treatments, N60 (60 kg per ha as urea split into two applications) and N180 (180 kg per ha in three splits) being applied  (Table \@ref(tab:N-rates)).

@@

As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using Bayesian multivariate generalised linear mixed models implemented in the R package `MCMCglmm` [@MCMCglmm2010]. Six models were created where the the dependent variables were 2015 tiller ShB incidence, 2015 tiller ShB severity, 2015 leaf ShB severity; 2016 tiller ShB incidence, 2016 tiller ShB severity; 2016 leaf ShB severity, respectively. The independent variables were the irrigation management and nitrogen rate and replicate was treated as a random variable. The base levels (control treatments) for the analyses for 2015 were: N rate, 0 kg/ha; irrigation management, FLD and for 2016 were: N rate, 60 kg/ha; irrigation management, FLD. Four MCMC chains were run for 55,000 steps, with the first 5000 discarded as burn in. The remaining 50,000 samples were thinned by taking every tenth sample resulting in 5,000 independent draws from the posterior distribution of the parameters of the model. The `coda` [@coda2006] package was used to provide utilities to check convergence using visual inspection of graphs and the Gelman-Rubin test. All data generated from this project and code used for analysis are available as a reproducible research compendium from <https://github.com/openplantpathology/rice_awd_pests> [@Sparks2018].

@@

In 2015 the incidence of tiller ShB remained low throughout the growing season (Fig. \@ref(Fig.:ShB-W-progress)a, \@ref(Fig.:ShB-N-progress)a). Water management was not significantly different (Fig. \@ref(Fig.:ShB-W-box)a, \@ref(Fig.:estimates-plot)a). However, the nitrogen treatments, N100 and N120, were both significantly different when compared with the control N0 treatment (Fig. \@ref(Fig.:ShB-N-box)a, \@ref(Fig.:estimates-plot)a).

@@

The changes to the inoculation methodology resulted in a higher rate of infection in 2016, with the N180 treatment reaching a maximum value of 98% incidence at the third observation (Fig. \@ref(Fig.:ShB-N-progress)b). The nitrogen treatment N180, was significantly different than the control N60 treatment (Fig. \@ref(Fig.:ShB-N-box)b, \@ref(Fig.:estimates-plot)b). As in the 2015 study, water management did not significantly differ \@ref(Fig.:ShB-W-box)b, \@ref(Fig.:estimates-plot)b).

@@

In 2016 the neither of the treatments, nitrogen rate or water management, were significantly different from the control treatments for leaf ShB severity (Fig. \@ref(Fig.:ShB-W-box)f, \@ref(Fig.:estimates-plot)f, \@ref(Fig.:ShB-N-box)f \@ref(Fig.:estimates-plot)f).

>>>>

Sheath blight (ShB) disease (*Rhizoctonia solani* Kühn), anastamosis group 1 [(teleomorph: *Thanatephorus cucumeris* (A.B. Frank) Donk.] of rice is an economically important disease worldwide throughout tropical and temperate production areas. Under conducive conditions he disease can cause up to 50\% yield losses [@Marchetti1991]. Studies by Savary et al. [-@Savary2000] estimated normal yield losses in tropical irrigated lowland rice in Asia at 5-10\%. The disease is best managed through an integrated disease management approach which includes resistance [@Kumar2009, @Srinivasachary2011] fungicide applications where affordable [@Rush1983; @Groth2008], nitrogen (N) management [@Slaton2003; @Castilla1996; @Tang2007] and water management [@Castilla1996].

@@

The plot design was a split-plot design with four replicates where irrigation was the main plot and N rate was the split-plot.

@@

The split-plot treatment rates of N were N0 (no N supply), N100 (100 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation) and N120 (120 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation) (Table \@ref(tab:N-rates)).

@@

Split-plot N rates differed from the 2015 rates with only two N treatments, N60 (60 kg per ha as urea split into two applications) and N180 (180 kg per ha in three splits) being applied  (Table \@ref(tab:N-rates)).

@@

As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using Bayesian multivariate generalised linear mixed models implemented in the R package `MCMCglmm` [@MCMCglmm2010]. Six models were created where the the dependent variables were 2015 tiller ShB incidence, 2015 tiller ShB severity, 2015 leaf ShB severity; 2016 tiller ShB incidence, 2016 tiller ShB severity; 2016 leaf ShB severity, respectively. The independent variables were the irrigation management and N rate and replicate was treated as a random variable. The base levels (control treatments) for the analyses for 2015 were: N rate, 0 kg/ha; irrigation management, FLD and for 2016 were: N rate, 60 kg/ha; irrigation management, FLD. Four MCMC chains were run for 55,000 steps, with the first 5000 discarded as burn in. The remaining 50,000 samples were thinned by taking every tenth sample resulting in 5,000 independent draws from the posterior distribution of the parameters of the model. The `coda` [@coda2006] package was used to provide utilities to check convergence using visual inspection of graphs and the Gelman-Rubin test. All data generated from this project and code used for analysis are available as a reproducible research compendium from <https://github.com/openplantpathology/rice_awd_pests> [@Sparks2018].

@@

In 2015 the incidence of tiller ShB remained low throughout the growing season (Fig. \@ref(Fig.:ShB-W-progress)a, \@ref(Fig.:ShB-N-progress)a). Water management was not significantly different (Fig. \@ref(Fig.:ShB-W-box)a, \@ref(Fig.:estimates-plot)a). However, the N treatments, N100 and N120, were both significantly different when compared with the control N0 treatment (Fig. \@ref(Fig.:ShB-N-box)a, \@ref(Fig.:estimates-plot)a).

@@

The changes to the inoculation methodology resulted in a higher rate of infection in 2016, with the N180 treatment reaching a maximum value of 98% incidence at the third observation (Fig. \@ref(Fig.:ShB-N-progress)b). The N treatment N180, was significantly different than the control N60 treatment (Fig. \@ref(Fig.:ShB-N-box)b, \@ref(Fig.:estimates-plot)b). As in the 2015 study, water management did not significantly differ \@ref(Fig.:ShB-W-box)b, \@ref(Fig.:estimates-plot)b).

@@

In 2016 the neither of the treatments, N rate or water management, were significantly different from the control treatments for leaf ShB severity (Fig. \@ref(Fig.:ShB-W-box)f, \@ref(Fig.:estimates-plot)f, \@ref(Fig.:ShB-N-box)f \@ref(Fig.:estimates-plot)f).
